Poincar\'e--Friedrichs constants over triangulated domains

### 2024-05-01 main.tex
more material on poincare inequality

% either proceeding simplex by simplex, matching along faces 
% or proceeding patch by patch, matching over intersections

>>>>

Let us assume that $p \in [1,\infty]$ and that $\Omega \subseteq \bbR^{n}$ is an open set. 
We say that a domain $\Omega \subseteq \bbR^{n}$ satisfies the Poincar\'e--Friedrichs inequality with exponent $p \in [1,\infty]$
if there exists a constant $C_{{\rm PF},\Omega,p} \geq 0$ such that the following holds:
for every vector field $\bff \in \nabla W^{1,p}(\Omega)$ there exists $u \in W^{1,p}(\Omega)$
such that $\nabla u = \bff$ and 
\begin{align}\label{math:inequalitypf:gradient}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm PF},\nabla,\Omega,p} 
    \| \bff \|_{L^{p}(\Omega)}
    .
\end{align}
We call $C_{{\rm PF},\Omega,p}$ the Poincar\'e--Friedrichs constant. 
This inequality implies an upper bound for the generalized inverse of the gradient operator $\nabla : W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)^{n}$.

\begin{remark}
    Before we proceed, we wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known are Poincar\'e inequality or also Friedrichs inequality in the literature.

    Given $p \in [1,\infty]$ and an open set $\Omega \subseteq \bbR^{n}$ of finite measure, 
    we say that $\Omega$ satisfies the Poincar\'e inequality with exponent $p$ 
    if there exists $C_{{\varnothing},\Omega,p} \geq 0$ such that 
    for all $u \in W^{1,p}(\Omega)$
    \begin{align}
        \| u - u_{\Omega} \|_{L^{p}(\Omega)}
        \leq 
        C_{\varnothing,\Omega,p} 
        \| \nabla u \|_{L^{p}(\Omega)}
        ,
    \end{align}
    where $u_{\Omega}$ is the \emph{average} of $u$ over $\Omega$, that is,
    \begin{align*}
        u_{\Omega} := |\Omega|^{-1} \int_{\Omega} u.
    \end{align*}
    Clearly, this Poincar\'e inequality implies the Poincar\'e--Friedrichs inequality and we have $C_{{\rm{PF}},\Omega,p} \leq C_{\varnothing,\Omega,p}$. 
    Towards a converse inequality, 
    let us first observe that the average of any $v \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
    \begin{align*}
        \| v_\Omega \|_{L^{p}(\Omega)}^{p}
        = 
        \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |v(x)| \right)^{p}
        \leq 
        \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |v(x)|^{p} \right)
        = 
        \| v \|_{L^{p}(\Omega)}^{p}
        .
    \end{align*}
    Here, we have used Jensen's inequality. 
    In the case $p = \infty$, any $v \in L^{\infty}(\Omega)$ satisfies $\| v_\Omega \|_{L^{\infty}(\Omega)} \leq \| v \|_{L^{\infty}(\Omega)}$. 
    We therefore combine 
    \begin{align*}
        \| v - v_\Omega \|_{L^{p}(\Omega)} 
        \leq
        2
        \| v \|_{L^{p}(\Omega)},
        \quad 
        v \in L^{p}(\Omega)
        .
    \end{align*}
    \begin{align*}
        \color{Emerald}
        \| u - u_\Omega \|_{L^{p}(\Omega)}^{p}
        \leq 
        2^{p-1} 
        \| u \|_{L^{p}(\Omega)}^{p}
        +
        2^{p-1} 
        \| u_\Omega \|_{L^{p}(\Omega)}^{p}
        \leq 
        2^{p} 
        \| u \|_{L^{p}(\Omega)}^{p}
        .
    \end{align*}
    We conclude that the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with $C_{\varnothing,\Omega,p} \leq 2 C_{{\rm PF},\Omega,p}$. 
    In the special case $p=2$, we always have $\| u - u_\Omega \|_{L^{2}(\Omega)} \leq \| u \|_{L^{2}(\Omega)}$ for any $v \in L^{2}(\Omega)$,
    and therefore it even follows that $C_{\varnothing,\Omega,2} \leq C_{{\rm PF},\Omega,2}$. 
    \color{red} Question: can we modify the preceding discussion for general $p$ so that this special inequality for $p=2$ emerges as a limit case? Then we get stronger results for $p \approx 2$. Stern's paper? \color{black}

    The Poincar\'e--Friedrichs inequality, which we interpret as an upper bound for the operator norm of a generalized inverse of the gradient is more natural for our discussion because later sections will develop analogous inequalities for the curl and the divergence operators. 
    The kernel of those operators is infinite-dimensional, whereas the gradient only has a one-dimensional kernel. 
    When discussing Banach spaces, there is no canonical orthogonal projection on the kernel, unlike in the Hilbert space case, 
    which is why the Poincar\'e--Friedrichs inequality seems to be the more natural approach.
    \color{red} Make this last paragraph for concise.
\end{remark}

% either proceeding simplex by simplex, matching along faces 
% or proceeding patch by patch, matching over intersections
\begin{theorem}
    \color{red}Precise form of the inequality yet to be discussed, though the construction is clear.
\end{theorem}

\begin{proof}
 Let $u \in W^{1,p}(\Omega)$. 
 For each $F \in \calT$ of dimension $n-1$, 
 we let $u_F \in W^{1,p}(U_F)$ such that $\nabla u_F = \nabla u$ over $U_F$ and such that $\| u_F \|_{L^{p}(U_F)} \leq C_{{\rm{P}},F} \| \nabla u_F \|_{L^{p}(U_F)}$.

 Since $\calT$ is strongly connected, there exists an enumeration of $n$-simplices $T_0, T_1, \dots, T_M$ such that for any $1 \leq m \leq M$ there exists some fixed but arbitrary $0 \leq j < m$ such that $F_m := T_m \cap T_j$ is a face of dimension $n-1$. We write $\Omega_m := \cup_{j=0}^{m} T_m$. Hence $\Omega = \Omega_M$.

%  We let $T \in \calT$ be the first $n$-simplex. 
%  There exists We already have

 Suppose that $1 \leq m \leq M$ such that there exists $w_{m-1} \in W^{1,p}(\Omega_{m-1})$ with $\nabla w_{m-1} = \nabla u$ over $\Omega_m$. Hence $c_{\star} := w_{m-1} - u_{|\Omega_{m-1}}$ is a constant. 
 By assumption, $T_{m}$ shares the face $F_{m}$ with $T_{j}$. 
 Since $\nabla u_{F_{m}|T_j} = \nabla w_{m-1|T_j}$,
 there exists a constant $c_{m} := w_{m-1|T_j} - u_{F_{m}|T_j}$.
 We define $u'_{F_m} := u_{F_m} + c_{m} \in W^{1,p}(U_F)$.
 By construction, $\nabla u'_{F_m} = \nabla u_{F_m} = \nabla u_{|U_{F_m}}$ and $u_{F_{m}|T_j} = w_{m-1|T_j}$. 
 But that already implies $u'_{F_m} = u_{|F_m} - c_{\star}$. 
 As we define a measurable function $u'_{\star} \in W^{1,p}(\Omega_m)$ by extending $w_{m-1}$ with $u'_{F_m|T_m}$.
 Moreover, 
 \begin{gather*}
    \| u'_{F_m} \|_{L^{p}(T_m)}
    \leq 
    \| u_{F_m} \|_{L^{p}(T_m)}
    +
    \| c_{m} \|_{L^{p}(T_m)},
    \qquad 
    \| c_{m} \|_{L^{p}(T_m)}
    \leq 
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_j)^{\frac 1 p} }
    \| c_{m} \|_{L^{p}(T_j)},
    \\ 
    \| c_{m} \|_{L^{p}(T_j)}
    \leq 
    \| w_{m-1}\|_{L^{p}(T_j)} + \| u_{F_{m}} \|_{L^{p}(T_j)} 
    .
 \end{gather*}
 In particular,
 \begin{align*}
    \| w_{m} \|_{L^{p}(T_m)}
    \leq 
    \| u_{F_m} \|_{L^{p}(T_m)}
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_j)^{\frac 1 p} }
    \| u_{F_{m}} \|_{L^{p}(T_j)}
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_j)^{\frac 1 p} }
    \| w_{m-1}\|_{L^{p}(T_j)}
    .
 \end{align*}
 In the special case $p < \infty$, we also find 
 \begin{align*}
    \| w_{m} \|_{L^{p}(T_m)}^{p}
    &\leq 
    3^{p-1}
    \| u_{F_m} \|_{L^{p}(T_m)}^{p}
    +
    3^{p-1}
    \frac{ \vol(T_m) }{ \vol(T_j) }
    \| u_{F_{m}} \|_{L^{p}(T_j)}^{p}
    +
    3^{p-1}
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_j)^{\frac 1 p} }
    \| w_{m-1}\|_{L^{p}(T_j)}^{p}
    \\&\leq 
    3^{p-1}
    \max\left( 1, \frac{ \vol(T_m) }{ \vol(T_j) } \right)
    \| u_{F_m} \|_{L^{p}(U_{F_m})}^{p}
    +
    3^{p-1}
    \frac{ \vol(T_m) }{ \vol(T_j) }
    \| w_{m-1}\|_{L^{p}(T_j)}^{p}
    .
 \end{align*}
 Recursive application provides the desired result. 
\end{proof}

\begin{remark}
    The previous result depends on only a few parameters of the triangulation:
    the length of any traversal from the root simplex,
    the ratios of the volumes of adjacent simplices,
    the Poincar\'e constants on each face patch, and the Poincar\'e constant of the initial simplex. 
    The aforementioned local Poincar\'e constants can generally be estimated in terms of shape-parameters. 
    If the face patches of the triangulation are convex, then better estimates are possible. 
\end{remark}
